**Density Dependence**

### 2023-02-13 Rmd/condMac.Rmd
clean

---
title: "**Nort East Atlantic Mackerel**"
subtitle: "Emprirical evidence for density dependence"
author: "L Kell & M Pastoors"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
mathjax: TRUE

@@

theme_set(theme_bw(16))

source("../R/get_dropbox.R")

matFn<-function(wt,x) 
   1/(1+exp(-x[1]*(wt-x[2])))

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"

```

@@

```{r}
dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
dat$age=factor(dat$age)

fn<-function(x,dat) 
   sum((dat$mat-matFn(dat$wt,x))^2)

par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)

ggplot(dat,aes(wt,mat))+
  geom_point(aes(wt,mat,col=age))+
  geom_smooth(se=F,span=0.2)+
  xlab("Mass-at-age")+ylab("Matutrity-at-age")+
  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
```

**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)

>>>>

---
title: "**Nort East Atlantic Mackerel**"
subtitle: "Mackerel: Emprirical evidence for density dependence"
author: "L Kell & M Pastoors"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
mathjax: TRUE

@@

theme_set(theme_bw(16))
```

```{r, dropbox}
source("../R/get_dropbox.R")

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

@@

```{r}
matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))

dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
dat$age=factor(dat$age)

fn<-function(x,dat) {
  sum((dat$mat-matFn(dat$wt,x))^2)}

matPar=optim(par=c(k=0.5,w50=.2), fn,dat=dat)

ggplot(dat,aes(wt,mat))+
  geom_point(aes(wt,mat,col=age))+
  geom_smooth(se=F,span=0.2)+
  xlab("Mass-at-age")+ylab("Matutrity-at-age")+
  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,matPar$par),wt=dat$wt))
```

**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`)

### 2023-02-13 Rmd/condWhb.Rmd
clean

source("../R/get_dropbox.R")
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"

matFn<-function(wt,x) 
   1/(1+exp(-x[1]*(wt-x[2])))

```

@@

```{r}
dat=model.frame(FLQuants(whb,"wt"=stock.wt,"mat"=function(x) mat(x)))
dat$age=factor(dat$age)

fn<-function(x,dat) 
   sum((dat$mat-matFn(dat$wt,x))^2)

>>>>

source("../R/get_dropbox.R")
dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

@@

```{r}
matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))

dat=model.frame(FLQuants(whb,"wt"=stock.wt,"mat"=function(x) mat(x)))
dat$age=factor(dat$age)

fn<-function(x,dat) {
  sum((dat$mat-matFn(dat$wt,x))^2)}

### 2023-02-13 Rmd/ddMac.Rmd
clean

@@

```{r}
parMat=FLPar(c(k=0.1, w5=par["a"]*par["l50"]^par["b"]))
dat=transform(dat,mat=1/(1+exp(-parMat[1]*(mass-parMat[2]))))

@@

    mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
  x=brp(x)}

@@

  mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
  m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
  x=brp(x)}

@@

```{r, check1}
par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","biomass"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))

f=FLQuant(c(rep(seq(0,1,length.out=51),each=101),rep(seq(1,25,length.out=61)[-1],each=101)),
          dimnames=list(year=dimnames(om)$year,iter=seq(101)))%*%refpts(eq)["msy","harvest"]
fbar(eq)=f

@@

om2=om

for (year in dimnames(om2)$year[-1]){
  om2=ddFn(year,om2,par)
  om2=fwd(om2,f=f[,year],sr=eq)
  }
catch(om2)=computeCatch(om2,slot="all")
```

@@

>>>>

```{r, dd-par}
par=rbind(par,
          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
                  bref =c(refpts(eq)["msy","biomass"]),
                  delta=-0.2,
                  matk = 0.2,
                  w50  =par["a"]*par["l50"]^par["b"],
                  alpha=-0.2)))
```

@@

```{r}
dat=transform(dat,mat=1/(1+exp(-par["matk"]*(mass-par["w50"]))))

@@

    mat(     x)=1/(1+exp(-par["matk"]*(stock.wt(x)-par["w50"])))
  x=brp(x)}

@@

lsr=landings.wt(eq)/stock.wt(eq)
dsr=discards.wt(eq)/stock.wt(eq)

# iterate
for (i in seq(10)){
  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],alpha=-0.2)
  landings.wt(x)=stock.wt(x)*lsr
  discards.wt(x)=stock.wt(x)*dsr

  mat(     x)=1/(1+exp(-par["matk"]*(stock.wt(x)-par["w50"])))
  m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
  x=brp(x)}

@@

```{r, check1}
f=FLQuant(c(rep(seq(0,1,length.out=51),each=101),rep(seq(1,25,length.out=61)[-1],each=101)),
          dimnames=list(year=dimnames(om)$year,iter=seq(101)))%*%refpts(eq)["msy","harvest"]
fbar(eq)=f

@@

om2=om
for (year in dimnames(om2)$year[-1]){
  om2=ddFn(year,om2,par)
  om2=fwd(om2,f=f[,year],sr=eq)
  }
catch(om2)=computeCatch(om2,slot="all")
```

@@

```{r}
save(par,file="../data/om/parMac.RData")
```

### 2023-02-13 Rmd/ddWhb.Rmd
clean

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/whb/dd/",
               fig.path  ="../tex/figs/whbdata/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

@@

>>>>

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/whb/dd/",
               fig.path  ="../tex/figs/whb/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

```{r, dd-par}
par=rbind(par,
          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
                  bref =c(refpts(eq)["msy","biomass"]),
                  delta=-0.2,
                  matk = 0.1,
                  w50  =par["a"]*par["l50"]^par["b"],
                  alpha=-0.2)))
```

@@

```{r}
save(par,file="../data/om/parWhb.RData")
```

### 2023-02-13 Rmd/omWhb.Rmd
clean

---
title: "**Density Dependence**"
subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

@@

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mac/om/",
               fig.path  ="../tex/figs/mac/om/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

```{r}
load(file.path(dropboxdir,"data/om/mac.RData"))
```

```{r, ices}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)

@@

```{r, vpa}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)
vpa =ices+FLAssess:::VPA(ices)

@@

F  =propagate(window(fbar(vpa),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

@@

```{r, vpaM}
load("../data/om/parMac.RData")
par["m1"]  =par["m1"]/10
par["w50"] =par["w50"]/1000
par["matk"]=30

load(file.path(dropboxdir,"data/inputs/ices.RData"))
vpaM=ices[["mac.27.nea"]]
vpaM=window(vpaM,start=1991)
m(vpaM)=par["m1"]%*%(stock.wt(vpaM)%^%par["m2"])
vpaM=vpaM+FLAssess:::VPA(vpaM)

@@

F  =propagate(window(fbar(vpaM),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.5,length.out=51)[-1]),each=dim(F)[2])

@@

##################################
par["bref"]=subset(d1,Yield==max(Yield))[,"Biomass"]
##################################

@@

for (i in ac(2020:2050)) {
  vpaDDM =ddFn(i,vpaDDM,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDM =fwd(vpaDDM,control=control,sr=eq)}

@@

for (i in ac(2020:2050)) {
  vpaDDMM =ddFn(i,vpaDDMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMM =fwd(vpaDDMM,control=control,sr=eq)}

@@

for (i in ac(2020:2050)) {
  vpaDDMMM =ddFn(i,vpaDDMMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMMM =fwd(vpaDDMMM,control=control,sr=eq)}

>>>>

---
title: "**Density Dependence**"
subtitle: "Blue whiting: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

@@

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/whb/om/",
               fig.path  ="../tex/figs/whb/om/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

```{r}
load(file.path(dropboxdir,"data/om/whb.RData"))
```

```{r, ices}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["whb.27.1-91214"]]
ices=window(ices,start=1991)

@@

```{r, vpa}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["whb.27.1-91214"]]
ices=window(ices,start=1991)
vpa =ices+FLAssess:::VPA(ices)

@@

F  =propagate(window(fbar(vpa),start=2020),101)
F[]=rep(c(seq(0,                             c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy","harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

@@

```{r, vpaM}
load("../data/om/parWhb.RData")
par["m1"]  =par["m1"]/10
par["w50"] =par["w50"]/100
par["w50"] =0.08
par["matk"]=100

load(file.path(dropboxdir,"data/inputs/ices.RData"))
vpaM=ices[["whb.27.1-91214"]]
vpaM=window(vpaM,start=1991)
m(vpaM)=par["m1"]%*%(stock.wt(vpaM)%^%par["m2"])
vpaM=vpaM+FLAssess:::VPA(vpaM)

@@

F  =propagate(window(fbar(vpaM),start=2020),101)
F[]=rep(c(seq(0,                             c(refpts(eq)["msy",  "harvest"]),  length.out=51),
          seq(c(refpts(eq)["msy","harvest"]),c(refpts(eq)["crash","harvest"])*2,length.out=51)[-1]),each=dim(F)[2])

@@

####################################################
par["bref"]=subset(d1,Yield==max(Yield))[,"Biomass"]
####################################################

@@

for (i in ac(2020:2050)) {
  vpaDDM =ddFn(i,vpaDDM,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
  mat(vpaDDM)[1]=0
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDM =fwd(vpaDDM,control=control,sr=eq)}

@@

for (i in ac(2020:2050)) {
  vpaDDMM =ddFn(i,vpaDDMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
  mat(vpaDDMM)[1]=0
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMM =fwd(vpaDDMM,control=control,sr=eq)}

@@

for (i in ac(2020:2050)) {
  vpaDDMMM =ddFn(i,vpaDDMMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
  mat(vpaDDMMM)[1]=0
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMMM =fwd(vpaDDMMM,control=control,sr=eq)}
